Sparse Function-space Representation \\ of Neural Networks

### 2023-05-16 paper/main.tex
added dual conditionign paragraph
updated cl notation

\section{\our: Sparse function-space represenation of neural networks}
%\paragraph{GP in the dual parameters}
The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process $f_{\textrm{post}}$ that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
%
\begin{equation}  \label{eq:gp_pred}
  \myexpect_{p(f_i \mid\vy)}[f_i]= \vk_{\vx i}^\top \valpha \quad \text{and} \quad
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] = k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
%
where the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters are the vectors $\valpha, \vbeta \in \R^{N}$, defined as:\looseness-1
%
\begin{equation}
  \label{eq:dual_param}
  \alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f_i}\log p(y_i \mid f_i)]
  \quad \text{and} \quad
  \beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f_i f_i}\log p(y_i \mid f_i)] .
\end{equation}
%
The relationships specified above are valid for generic likelihoods, and involve no approximations since the expectation is under the exact posterior. This higlights the fact that our approximate inference technique that will follow, while usually viewed as a posterior approximation, can be alternatively interpreted as an approximation of the expectation of gradients of loss/likelihoods.

@@

\paragraph{\our for continual learning}
% Definition of CL -> a sequence of tasks 
In the continual learning setting, the training is divided into $T$ tasks, each with its training data set $\dataset_t = \{(\mathbf{x}_{i}, \mathbf{y}_{i})\}_{i=1}^{N_t}$, which after the task is discarded and cannot be accessed in the following stages of the training. Rehearsal and function regularization-based models keep a subset of the training data for each task to help the model alleviate forgetting. In the same way, after each task, we can use our \our method to build a compact representation of the neural network and then exploit this representation to build a functional regularizer to add to the training loss for the subsequent tasks. Recently proposed GP-based regularizers \cite{ pan2020continual, rudner2022continual} have shown better performances than weight space equivalents getting state-of-the-art on CL benchmarks. However, they are limited to use subset approximations of the posterior \cref{app:cl-experiment}.
%, which in practice means randomly selecting some data computing its posterior at only those points and then using that as the regularizer. 
A more principled approach would be to sparsify the resultant GP into a low-rank approximation to ensure information coming from the entire task dataset is not lost. %, essentially what the sparse dual variables do.   \todo{this term is replacing the missing data}
% Where the regularizer comes from
At the end of the training for each task $t$, we compute the Laplace-GGN approximated kernel function for the current model and randomly select $M$ inducing points. Given the issues described in the previous sections, related to computing the full covariance posterior $\MKxx$ for $\dataset_t$,\todo{this should be $\kappa(\MX_t, \MX_t)$} we can exploit the dual parameterization to efficiently encode the information from  $\dataset_t$ using the dual parameterization, as follows, 
%
\begin{equation}\textstyle
 	\bar{\MB}^{-1}_t = \MKzz^{-1} \MBeta_\vu \MKzz^{-1} \in \R^{M \times M}, 
 	\quad \text{s.t.} \quad
 	\MBeta_{\vu} =  \sum_{i \in \dataset_t} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
 	% \quad \forall t \in [1, T]
\end{equation}
%
where $\MKzz$ and $\vkzi$ are the gram matrix and vector computed on the inducing points $\MZ_t = \{\vz_i\}_{i=1}^M$ for task $t$ .\
%
We keep a memory buffer $\mathcal{M} = \{(\MZ_t, \vu_t, \bar{\MB}^{-1}_t)\}_{t=1}^T$, where $\MZ_t$ are the inducing points selected from $\dataset_t$ and  $\vu_t = f_{\vw_t^*}(\MZ_t)\in \R^{M}$ are the neural network outputs associated to them, computed with the MAP estimate $\weights_t^*$ after task $t$. Note that, for notational convenience, we are extending our neural network notation to be compatible with multiple inputs points, \ie, each $i$\textsuperscript{th} entry is $f_{\weights_t^*}(\vz_i)\in \R$. Moreover, we restrict ourselves to single-output neural networks and refer the reader to further details on the multi-output setting to \cref{sec:cl_multioutput}.

After each task, we update $\mathcal{M}$, and we can construct a regularizer term, which aims to reduce the drift of the current model from diverging on the inducing points $\MZ_s$ for all previous tasks $ s = 1, \ldots, t-1$. This means that for training  the successive task, the new objective to minimize becomes $\weights_t^* = \arg\min_{\weights} \mathcal{L}(\mathcal{D}_t, \weights) + \tau \, \mathcal{R}_\textrm{\our}(\weights, \mathcal{M})$, with the regularizer
%
\begin{equation}\label{eq-cl-regularizer}\textstyle
  \mathcal{R}_\mathrm{\our}(\weights, \mathcal{M}) = \frac{1}{2} \sum_{s=1}^{t-1} \frac{1}{M} 
	\left\lVert 
	f_{\weights}(\MZ_{s}) - \vu_t % f_{\weights_{s}}(\MZ_s)
	\right\rVert_{\bar{\MB}^{-1}_{s}}.
\end{equation}

>>>>

\section{\our: Sparse function-space represenation of neural networks}
%\paragraph{GP in the dual parameters}
The seminal work by \citet{csato2002sparse} (parameterization Lemma~1) gives a parameterization for the posterior process $f_{\textrm{post}}$ that can be found through the Bayesian update using the GP prior and the likelihood function. This gives a `dual' parameterization, $\valpha$ and $\vbeta$,
%
\begin{equation}  \label{eq:gp_pred}
  \myexpect_{p(f_i \mid\vy)}[f_i]= \vk_{\vx i}^\top \valpha \quad \text{and} \quad
  \mathrm{Var}_{p(f_i \mid \vy)}[f_i] = k_{ii} - \vk_{\vx i}^\top ( \MKxx + \diag(\vbeta)^{-1})^{-1} \vk_{\vx i},
\end{equation}
%
where the $ij$\textsuperscript{th} entry of the matrix $\MKxx \in \R^{N \times N}$ is $\kappa(\vx_i,\vx_j)$, $\vk_{\vx i}$ denotes a vector where each $j$\textsuperscript{th} element is $\kappa(\vx_i, \vx_j)$, and $k_{ii} = \kappa(\vx_i, \vx_i)$.  The dual parameters are the vectors $\valpha, \vbeta \in \R^{N}$, defined as:\looseness-1
%
\begin{equation}
  \label{eq:dual_param}
  \alpha_i \coloneqq \myexpect_{p(\vw \mid \vy)}[\nabla_{f}\log p(y_i \mid f) |_{f=f_i}]
  \quad \text{and} \quad
  \beta_i \coloneqq - \myexpect_{p(\vw \mid \vy)}[\nabla^2_{f f}\log p(y_i \mid f_i) |_{f=f_i}],
\end{equation}
%
where $f_\vw(\vx_i) = f_i$ above. The relationships specified above are valid for generic likelihoods, and involve no approximations since the expectation is under the exact posterior. This higlights the fact that our approximate inference technique that will follow, while usually viewed as a posterior approximation, can be alternatively interpreted as an approximation of the expectation of gradients of loss/likelihoods.

@@

\paragraph{\our for continual learning}
% Definition of CL -> a sequence of tasks 
In the continual learning setting, the training is divided into $T$ tasks, each with its training data set $\dataset_t = \{(\mathbf{x}_{i}, \mathbf{y}_{i})\}_{i=1}^{N_t}$, which after the task is discarded and cannot be accessed in the following stages of the training. Rehearsal and function regularization-based models keep a subset of the training data for each task to help the model alleviate forgetting. In the same way, after each task, we can use our \our method to build a compact representation of the neural network and then exploit this representation to build a functional regularizer to add to the training loss for the subsequent tasks. Recently proposed GP-based regularizers \cite{ pan2020continual, rudner2022continual} have shown better performances than weight space equivalents getting state-of-the-art on CL benchmarks. However, they are limited to use subset approximations of the posterior \cref{app:cl-experiment}.
%, which in practice means randomly selecting some data computing its posterior at only those points and then using that as the regularizer. 
A more principled approach would be to sparsify the resultant GP into a low-rank approximation to ensure information coming from the entire task dataset is not lost. %, essentially what the sparse dual variables do.   \todo{this term is replacing the missing data}
% Where the regularizer comes from
At the end of the training for each task $t$, we compute the Laplace-GGN approximated kernel function for the current model and randomly select $M$ inducing points. Given the issues described in the previous sections, related to computing the full covariance posterior $\MKxx$ for $\dataset_t$,\todo{this should be $\kappa(\MX_t, \MX_t)$} we can exploit the dual parameterization to efficiently encode the information from  $\dataset_t$ using the dual parameterization, as follows, 
%
\begin{equation}\textstyle
 	\bar{\MB}^{-1}_t = \MKzz^{-1} \MBeta_\vu \MKzz^{-1} \in \R^{M \times M}, 
 	\quad \text{s.t.} \quad
 	\MBeta_{\vu} =  \sum_{i \in \dataset_t} \vkzi \,\hat{\beta}_{i} \, \vkzi^{\T} ,    
 	% \quad \forall t \in [1, T]
\end{equation}
%
where $\MKzz$ and $\vkzi$ are the gram matrix and vector computed on the inducing points $\mathcal{Z}_t = \{\vz_i\}_{i=1}^M$ selected from $\dataset_t$.\
%
We keep a memory buffer $\mathcal{M} = \{(\mathcal{Z}_t, \vu_t, \bar{\MB}^{-1}_t)\}_{t=1}^T$, where $\vu_t \in \R^{M}$ is a vector with entries $u_{t, i} = f_{\vw_t^*}(\vz_i), \forall \vz_i \in \mathcal{Z}_t$ are the neural network outputs computed on the inducing points with the MAP estimate $\weights_t^*$ after task $t$. For notational convenience, we restrict here to single-output neural networks and refer the reader to further details on the multi-output setting to \cref{sec:cl_multioutput}.

After each task, we update $\mathcal{M}$, and we can construct a regularizer term, which aims to reduce the drift of the current model from diverging on the inducing points $\MZ_s$ for all previous tasks $ s = 1, \ldots, t-1$. This means that for training  the successive task, the new objective to minimize becomes $\weights_t^* = \arg\min_{\weights} \mathcal{L}(\mathcal{D}_t, \weights) + \tau \, \mathcal{R}_\textrm{\our}(\weights, \mathcal{M})$, with weight $\tau$ and regularizer
%
\begin{equation}\label{eq-cl-regularizer}\textstyle
  \mathcal{R}_\mathrm{\our}(\weights, \mathcal{M}) = \frac{1}{2} \sum_{s=1}^{t-1} \frac{1}{M} 
	\left\lVert 
	\vu_t - \vu_s % f_{\weights_{s}}(\MZ_s)
	\right\rVert_{\bar{\MB}^{-1}_{s}}, \quad \vu_t \in \R^M, \text{with } u_{s,i} = f_{\weights}(\vz_{i}) \forall {\vz_i} \in \mathcal{Z}_s .
\end{equation}
